---
title: "ReproducableResearchCourseProject1"
author: "Wolfgang Brandl"
date: "18 Juni 2016"
output: html_document
---
```{r, echo=FALSE, results='hide', warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(data.table)
library(scales)
library(Hmisc)
```
## Set Work Area 
```{r, results='markup', warning=TRUE, message=TRUE}
wd <- file.path ("D:","Users","wbrandl","Coursera","ReproducableResearchCourseProject1")
setwd (wd)
if (!file.exists("data")){
  dir.create("data")
}
wd <- file.path ("D:","Users","wbrandl","Coursera","ReproducableResearchCourseProject1","data")
setwd (wd)
```

##Download and Loading and preprocessing the data
##### 1. Download data file

```{r}
download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip", "repdataFdataFactivity.zip", method="wininet", quiet = FALSE, mode = "wb",cacheOK = TRUE)
```
##### 1. Load the data (i.e. read.csv())
```{r, results='markup', warning=TRUE, message=TRUE}
data <- read.csv(unz("repdataFdataFactivity.zip", "activity.csv"), header=T, quote="\"", sep=",")
```
## What is mean total number of steps taken per day?
```{r}
library(data.table)
datatable <- data.table(data)
total.steps <- datatable[steps>0,.(stepsperday = sum(steps)),by=.(date)]
```
##### 1. Make a histogram of the total number of steps taken each day
```{r}
qplot(total.steps$stepsperday, binwidth=1000, xlab="total number of steps taken each day")
```
##### 2. Calculate and report the mean and median total number of steps taken per day
```{r}
mean(total.steps$stepsperday, na.rm=TRUE)
median(total.steps$stepsperday, na.rm=TRUE)
```

## What is the average daily activity pattern?
```{r}
daily.averages <- datatable[!is.na(steps),.(steps=as.integer(mean(steps))),by=.(interval)]
```

##### 1. Make a time series plot
```{r}
ggplot(data=daily.averages, aes(x=interval, y=steps)) +
  geom_line() +
  xlab("5-minute interval") +
  ylab("average number of steps taken")
```

##### 2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
daily.averages[which.max(daily.averages$steps),]
```

## Imputing missing values
##### 1. Calculate and report the total number of missing values in the dataset 

```{r how_many_missing}
missing <- is.na(data$steps)
# How many missing
table(missing)
```

##### 2. Devise a strategy for filling in all of the missing values in the dataset.
##### 3. Create a new dataset that is equal to the original dataset but with the missing data filled in.

```{r}
activityDataImputed <- data
activityDataImputed$steps <- activityDataImputed$steps <- impute(data$steps, fun=mean)
```
##### 4. Make a histogram of the total number of steps taken each day 
```{r}
stepsByDayImputed <- tapply(activityDataImputed$steps, activityDataImputed$date, sum)
qplot(stepsByDayImputed, xlab='Total steps per day (Imputed)', ylab='Frequency using binwith 500', binwidth=500)
```

##### ... and Calculate and report the mean and median total number of steps taken per day. 
```{r}
stepsByDayMeanImputed <- mean(stepsByDayImputed)
stepsByDayMedianImputed <- median(stepsByDayImputed)
```
* Mean (Imputed): `r stepsByDayMeanImputed`
* Median (Imputed):  `r stepsByDayMedianImputed`


----

## Are there differences in activity patterns between weekdays and weekends?
##### 1. Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.

```{r}
activityDataImputed$dateType <-  ifelse(as.POSIXlt(activityDataImputed$date)$wday %in% c(0,6), 'weekend', 'weekday')
```

##### 2. Make a panel plot containing a time series plot

```{r}
averagedActivityDataImputed <- aggregate(steps ~ interval + dateType, data=activityDataImputed, mean)
ggplot(averagedActivityDataImputed, aes(interval, steps)) + 
    geom_line() + 
    facet_grid(dateType ~ .) +
    xlab("5-minute interval") + 
    ylab("avarage number of steps")
```